\newpage
\phantomsection
\label{prf:metrics-are-pseudo-metrics}
\LRAProofFor{lem:metrics-are-pseudo-metrics}

\begin{remark*}[Return]
\hyperref[lem:metrics-are-pseudo-metrics]{Return to Lemma}
\end{remark*}

\begin{lemma*}[Metrics are Pseudo-Metrics]
Every metric on a nonempty set \(X\) is a pseudo-metric on \(X\).
The converse is false: there exist pseudo-metrics that are not metrics.
\end{lemma*}

\begin{proof}[Professional Standard Proof]
\LRAProofBodyStart
Let \(d\) be a metric on \(X\).  The metric axioms give
\[
d(x,y)\geq 0,\qquad d(x,y)=d(y,x),\qquad
d(x,y)\leq d(x,z)+d(z,y)
\]
for all \(x,y,z\in X\).  They also give \(d(x,x)=0\), since
\(d(x,y)=0\) if and only if \(x=y\).  Hence \(d\) satisfies every
pseudo-metric axiom.

For the converse, let \(X=\{a,b\}\) with \(a\neq b\), and define
\(d:X\times X\to\mathbb{R}\) by \(d(x,y)=0\) for all \(x,y\in X\).
Then \(d\) is nonnegative, symmetric, satisfies \(d(x,x)=0\), and
satisfies the triangle inequality.  Thus \(d\) is a pseudo-metric.
But \(d(a,b)=0\) while \(a\neq b\), so \(d\) is not a metric.
\end{proof}

\begin{proof}[Detailed Learning Proof]
\LRAProofBodyStart

\textbf{Step 1.} Compare the axioms.

A metric has nonnegativity, symmetry, and the triangle inequality.  These are
exactly three of the pseudo-metric requirements.  The remaining pseudo-metric
requirement is \(d(x,x)=0\), which follows from the metric condition
\[
  d(x,y)=0 \Longleftrightarrow x=y
\]
by substituting \(y=x\).  Therefore every metric is a pseudo-metric.

\textbf{Step 2.} Build a pseudo-metric that fails to be a metric.

Take a two-point set \(X=\{a,b\}\) and assign distance zero to every pair:
\[
  d(a,a)=d(a,b)=d(b,a)=d(b,b)=0.
\]
All pseudo-metric axioms hold.  Nonnegativity and symmetry are immediate, and
the triangle inequality says \(0\leq 0+0\), which is true.

\textbf{Step 3.} Identify the failed metric axiom.

The metric zero-distance condition would require
\[
  d(a,b)=0 \Longrightarrow a=b.
\]
But \(d(a,b)=0\) and \(a\neq b\).  Hence this pseudo-metric is not a metric,
so the converse implication is false.
\end{proof}

\begin{remark*}[Proof structure]
The forward direction is an axiom comparison.  The reverse direction is
disproved by a two-point zero-distance example.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:metric-space]{Definition: Metric Space}
  \item \hyperref[def:pseudo-metric]{Definition: Pseudo-Metric}
\end{itemize}
\end{dependencies}

\clearpage
