\chapter{Lattice Order}

\begin{tcolorbox}[
  colback=breadcrumb!8,
  colframe=breadcrumb!50,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{lattice-order}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Sets, Relations, and Functions
$\;\to\;$ \textbf{Lattice Order}
$\;\to\;$ Algebras of Sets
\end{center}
\end{tcolorbox}

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Roadmap}},
  fonttitle=\small\bfseries
]
\noindent\textbf{How we got here.}
The preceding chapter on sets, relations, and functions already introduced
partial orders, chains, suprema, infima, and the first appearance of
lattices as an example of order structure.  The material there was sufficient
for the purposes of that chapter.

\medskip
\noindent\textbf{What this chapter develops.}
Here order is the main object of study rather than an incidental feature.
The chapter builds a self-contained theory of ordered sets: the canonical
definitions of partial, strict, and total orders; chains and antichains;
maps between ordered sets; the duality principle; and the covering relation
and its role in finite posets.

\medskip
\noindent\textbf{Where this leads.}
Lattice order is the structural substrate beneath algebras of sets,
sigma-algebras, topologies, Boolean algebras, and fixed-point theorems.
The fuller development of distributive, modular, and complete lattices
belongs in a later chapter dedicated to that theory.
\end{tcolorbox}

\input{volume-iv/algebra/order-and-lattice/lattice-order/notes/index}

\section*{Proofs}
\input{volume-iv/algebra/order-and-lattice/lattice-order/proofs/index}
